\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-11-25 Limber_Comment.tex
First draft of photo-z section and conclusion
Full draft; TBD the theory section!

The measurement of weak gravitational lensing by large scale structures provides a powerful cosmological probe of dark matter, dark energy and modifications to gravity.  As such it is the primary science goal of several current (KiDS, HSC, DES) and future (Euclid, LSST, WFIRST) large-scale surveys.   Interest in the results from these surveys is high as statistically significant tensions have been found between the cosmological parameter constraints from the CMB Planck experiment \citep{planck/cosmo:2015} in comparison to weak lensing constraints from both the Kilo-Degree Survey (KiDS) and the Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS).  If the source of this tension is not a result of so-far unconsidered sources of systematic errors in one or all experiments, extensions to the standard flat $\Lambda$CDM cosmological need to be considered.  \citet{joudaki/etal:2017} have shown, for example, that the tension can be resolved with an evolving dark energy model.

In this comment paper we discuss four approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation function.   These approximations have been explored in (ADD REF) and most recently highlighted by \citet{kitching/etal:2016} as a way to reduce tensions between weak lensing and CMB experiments, focusing on a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.    In section~\ref{sec:cfhtlens} we compare the `one-parameter' cosmological re-analysis of \citet{kitching/etal:2016} with the full six-parameter analysis by \citet{kilbinger/etal:2013}, arguing that this particular non-tomographic data set is not in significant tension with Planck.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published.  In section~\ref{sec:theory} we clarify that only one/two? of the approximations highlighted by \citet{kitching/etal:2016} have actually been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017} and discuss why we do not expect these approximations to impact on the cosmological analysis of these surveys.    Finally in section~\ref{sec:photoz} we review our current knowledge of photometric redshift uncertainties and the sophisticated mitigation strategies that have already been applied to data from CFHTLenS, DES and KiDS.  We suggest that a possible misinterpretation in the direction of the redshift biases reported in \citet{choi/etal:2016} has lead to the incorrect conclusion of \citet{kitching/etal:2016} that these redshift biases can resolve the tension with Planck.

@@

\section{Conclusion}

>>>>

The measurement of weak gravitational lensing by large scale structures provides a powerful cosmological probe of dark matter, dark energy and modifications to gravity.  As such it is the primary science goal of several current (KiDS, HSC, DES) and future (Euclid, LSST, WFIRST) large-scale surveys.   Interest in the results from these surveys is high as statistically significant deviations have been found between the cosmological parameter constraints from the CMB Planck experiment \citep{planck/cosmo:2015} in comparison to weak lensing constraints from both the Kilo-Degree Survey \citep[KiDS;][]{hildebrandt/etal:2016} and the Canada-France-Hawaii Telescope Lensing Survey \citep[CFHTLenS;][] {joudaki/etal:2016}.  If the source of this tension is not a result of so-far unconsidered sources of systematic errors in one or all experiments, extensions to the standard flat $\Lambda$CDM cosmological need to be considered.  \citet{joudaki/etal:2017} have shown, for example, that the tension can be resolved with an evolving dark energy model.

In this comment paper we discuss four approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation function.   These approximations have been explored in (ADD REF) and most recently highlighted by \citet{kitching/etal:2016} as a way to reduce tensions between weak lensing and CMB experiments, focusing on a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.    In section~\ref{sec:cfhtlens} we compare the `one-parameter' cosmological re-analysis of \citet{kitching/etal:2016} with the full six-parameter analysis by \citet{kilbinger/etal:2013}, arguing that this particular non-tomographic data set is not in significant tension with Planck.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published.  In section~\ref{sec:theory} we clarify that only one/two? of the approximations highlighted by \citet{kitching/etal:2016} have actually been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017} and discuss why we do not expect these approximations to impact on the cosmological analysis of these surveys.    Finally in section~\ref{sec:photoz} we review our current knowledge of photometric redshift uncertainties and the sophisticated mitigation strategies that have already been applied to data.  We suggest that a possible misinterpretation in the direction of the redshift biases reported in \citet{choi/etal:2016} has lead to the incorrect conclusion of \citet{kitching/etal:2016} that these redshift biases can resolve the tension with Planck.

@@

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion.tex}

### 2016-11-25 cfhtlens.tex
Full draft; TBD the theory section!

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to Planck constraints.  They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard analysis' which we review in Section~\ref{sec:theory}.    We argue that this artificial reduction in the error could easily mislead as it makes it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.

We conclude this section by noting that all the advances in understanding listed above were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016}.  The CFHTLenS tomographic cosmological analysis was also revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis, as discussed in more detail in Section~\ref{sec:photoz}.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised measurements and calibrations for the shear and photometric redshifts.

>>>>

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to Planck constraints.  They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard analysis' which we review in Section~\ref{sec:theory}.    We argue that this artificial reduction in the error could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.

All the advances in understanding listed above were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports at $2.3 \sigma$ tension with Planck.  The CFHTLenS tomographic cosmological analysis was also revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis, as discussed in more detail in Section~\ref{sec:photoz}.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised measurements and calibrations for the shear and photometric redshifts.

### 2016-11-25 conclusion.tex
Full draft; TBD the theory section!

In this comment we have highlighted a number of shortcomings in the analysis presented in \cite{kitching/etal:2016}.  We argue that the flat sky and Limber approximations that can be used in cosmic shear analyses are not able to explain away the tension between weak lensing and Planck results, in contrast to the message that \cite{kitching/etal:2016} convey.  Our critique should not, however, detract from the extremely important core message of their paper, that future surveys will need to carefully consider these approximations and optimise their statistical analyses accordingly.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic renders the cosmic shear measurement completely insensitive to the low-$\ell$ scales where these approximations become important \citep{schneider/etal:2010}.

>>>>

In this comment we have highlighted a number of shortcomings in the analysis presented in \cite{kitching/etal:2016}.  We argue that the flat sky and Limber approximations that can be used in cosmic shear analyses are not able to explain away the tension between current weak lensing and Planck results, in contrast to the message that \cite{kitching/etal:2016} convey.  Our critique should not, however, detract from the extremely important core message of their paper, that future surveys will need to carefully consider these approximations and optimise their statistical analyses accordingly.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic renders the cosmic shear measurement completely insensitive to the low-$\ell$ scales where these approximations become important \citep{schneider/etal:2010}.

### 2016-11-25 photoz.tex
Full draft; TBD the theory section!

In \citet{choi/etal:2016} a cross-correlation clustering analysis, between photometric redshift slices and overlapping spectroscopic redshifts, is used to determine linear biases in the CFHTLenS photometric redshift distributions.   The maximum redshift $z_B<0.9$ in this analysis was limited by the redshift overlap with the spectroscopic sample  \citet{choi/etal:2016} found significant biases and concluded that any re-analysis of CFHTLenS should include systematic error terms to account for bias and scatter.    This re-analysis was presented in \citet{joudaki/etal:2016} where the impact of correcting for the biases determined by \citet{choi/etal:2016}, including their associated errors, served to reduce the overall constraining power of the survey and hence also the tension between constraints.  The best-fit model, however moved further from the Planck-preferred model, in agreement with the `toy-model' analysis of \citet{choi/etal:2016} which concluded that the effect of these biases would be to {\it reduce} the recovered amplitude of $\sigma_8$ by $\sim 4$\%.

@@

\cite{kitching/etal:2016} use a linear relation fit to the biases determined by \citet{choi/etal:2016} in order to extrapolate the results beyond the photometric redshift limit of $z_B<0.9$ and determine a bias at each redshift.  They then correct the non-tomographic \citet{kilbinger/etal:2013} redshift distribution with this bias relation.  They report that this correction results in an {\it increase} to the recovered amplitude of $\sigma_8$ by $\sim 4$\%, which resolves the tension with Planck.   We are unable to recover this conclusion using two different methods.   Based on the \cite{kitching/etal:2016} relation, the predicted shift at the mean CFHTLenS redshift $z=0.747$ is $\Delta_z = 0.056$, where we use the convention from \citet{choi/etal:2016} that a positive $\Delta_z$ serves to increase the overall mean redshift.  Using the same relation to individually shift each component of the full \citet{kilbinger/etal:2013} redshift histogram also results in a change to the mean redshift of $\Delta_z = 0.056$.  Increasing the mean redshift implies a reduction in the recovered amplitude of $\sigma_8$.  Using the same toy-model' analysis of \citet{choi/etal:2016} we would expect this change in the mean redshift to {\it reduce} the recovered amplitude of $\sigma_8$ by $\sim 5$\%.    From this we can only conclude that \cite{kitching/etal:2016} have misinterpreted the direction of the redshift biases reported in \citet{choi/etal:2016}.

>>>>

In \citet{choi/etal:2016} a cross-correlation clustering analysis, between photometric redshift slices and overlapping spectroscopic redshifts, is used to determine linear biases in the CFHTLenS photometric redshift distributions.   The maximum redshift $z_B<0.9$ in this analysis was limited by the redshift overlap with the spectroscopic sample.  \citet{choi/etal:2016} found significant biases and concluded that any re-analysis of CFHTLenS should include systematic error terms to account for bias and scatter.    This re-analysis was presented in \citet{joudaki/etal:2016} where the impact of correcting for the biases determined by \citet{choi/etal:2016}, including their associated errors, served to reduce the overall constraining power of the survey and hence also the tension between constraints.  The best-fit model, however moved further from the Planck-preferred model, in agreement with the `toy-model' analysis of \citet{choi/etal:2016} which concluded that the effect of these biases would be to {\it reduce} the recovered amplitude of $\sigma_8$ by $\sim 4$\%.

@@

\cite{kitching/etal:2016} use a linear relation fit to the biases determined by \citet{choi/etal:2016} in order to extrapolate the results beyond the photometric redshift limit of $z_B<0.9$ and determine a bias at each redshift.  They then correct the non-tomographic \citet{kilbinger/etal:2013} redshift distribution with this bias relation.  They report that this correction results in an {\it increase} to the recovered amplitude of $\sigma_8$ by $\sim 4$\%, which resolves the tension with Planck.   We are unable to recover this conclusion using two different methods.   Based on the \cite{kitching/etal:2016} relation, the predicted shift at the mean CFHTLenS redshift $z=0.747$ is $\Delta_z = 0.056$, where we use the convention from \citet{choi/etal:2016} that a positive $\Delta_z$ serves to increase the true mean redshift.  Using the same relation to individually shift each component of the full \citet{kilbinger/etal:2013} redshift histogram also results in a change to the mean redshift of $\Delta_z = 0.056$.  Increasing the mean redshift implies a reduction in the recovered amplitude of $\sigma_8$.  Using the same toy-model' analysis of \citet{choi/etal:2016} we would expect this change in the mean redshift to {\it reduce} the recovered amplitude of $\sigma_8$ by $\sim 5$\%.    From this we suggest that \cite{kitching/etal:2016} could have misinterpreted the direction of the redshift biases reported in \citet{choi/etal:2016}.
